---
title: "Caravan Insurance Challenge"
author: "Yash"
date: "June 24, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#install.packages("rmarkdown")
library(randomForest)
library(plyr)
library(rmarkdown)
library(tidyverse)
library(sqldf)
library(reshape2)
library(scales)
#install.packages("ROSE")
library(ROSE)
#install.packages("caret")
#install.packages('e1071', dependencies=TRUE)
library(caret)
```

# Predicting Caravan Insurance Purchases

##  The data is provided by the Caravan Insurance company of its customer base. Some of the customers have already bought  policies for  fire risk , auto , boat etc from Caravan. Using this information we need predict whether they will purchase Caravan Home Insurance 

## This is a classification problem. It is a high dimension dataset with 86 variables. The number of people who have bought insurance is very low, less than 5%. We will use tree based methods like Random forest and Boosted trees to predict.


### Load the training data and inspect the data 

```{r message=FALSE}
Cins <-  read.csv.sql("C1.csv","select * from file where ORIGIN = 'train' ", header=TRUE, sep=",")

### Exclude the  first column 
Caravan_train <- Cins[-1]


### Check the dimensions
dim(Caravan_train )
#names(Caravan_train )


# Inspect the data 
str(Caravan_train )
#head(Caravan_train )

```
###  check for missing  values 

```{r}
Missing_Count <-  sum(is.na.data.frame(Caravan_train))
print(paste ("Missing count -> " , Missing_Count))
```

### Its an highly unbalanced dataset. This is evident from the below bar chart 

```{r}
bar_label <- factor(Caravan_train$Number_of_mobile_home_policies  , labels=c("NotPurchased","Purchased"))
#bar_label
ggplot(data =  Caravan_train, aes(y = "", x =Number_of_mobile_home_policies , fill = bar_label ) )+ 
  geom_bar(stat = "identity" ) + 
  ggtitle("Not Purchased /  Purchased")
print(paste("Insurance count : 0 -> not having insurance , 1 - Having Insurance"))
table(Caravan_train$Number_of_mobile_home_policies)

```
## Data Exploratory


###  Education levels of population who have bought insurance. People with low education level are more likey to buy 
```{r}
temp1 <-  ""
templ1 <-  ""
temp2 <- ""
grp_data <-  ""
grp_data <- sqldf(" select High_level_education	 , Medium_level_education , Lower_level_education   from Caravan_train where Number_of_mobile_home_policies  = 1  ")
class(grp_data)
dim(grp_data)
temp2 <- gather(grp_data,Education,Value,1:3)

ggplot(data= temp2, aes (x = Value,y = "",fill = Value)) + 
  geom_bar(stat = 'identity') +
  facet_grid(. ~ Education ) + 
  ggtitle("Education distribution")


```

### Distribution of insurance buyers by Car ownership. People with 1 car are more likely to buy 
```{r}

temp1 <-  ""
templ1 <-  ""
temp2 <- ""
grp_data <-  ""
grp_data <- sqldf(" select X1_car , X2_cars , No_car from Caravan_train where Number_of_mobile_home_policies  = 1  ")
class(grp_data)
dim(grp_data)
names(grp_data) <- c("1 CAR","2 CAR","0 CAR")


temp2 <- gather(grp_data,CarOwnership,Value,1:3)

ggplot(data= temp2, aes (x = Value,y = "",fill = Value)) + 
  geom_bar(stat = 'identity') +
  facet_grid(. ~ CarOwnership  ) + 
  ggtitle("CarOwnership")

``` 


### Insurance policy count.  We  can see corelation between people having car policies3 party policy and fire risk policies.

```{r}
  temp1 <-  ""
templ1 <-  ""
temp2 <- ""
grp_data <-  ""
grp_data <- sqldf(" select 
				   private_third_party_insurance                
, third_party_insurance_firms                  
, third_party_insurane_agriculture             
, car_policies                                 
, delivery_van_policies                        
, Contribution_motorcycle_scooter              
, lorry_policies                               
, trailer_policies                             
, tractor_policies                             
, agricultural_machines_policies               
, moped_policies                               
, life_insurances                              
, private_accident_insurance_policies          
, family_accidents_insurance_policies          
, disability_insurance_policies                
, fire_policies                                
, surfboard_policies                           
, boat_policies                                
, bicycle_policies                             
, property_insurance_policies                  
, social_security_insurance_policies           

 from Caravan_train where Number_of_mobile_home_policies  = 1  ")
class(grp_data)
dim(grp_data)

temp2 <- gather(grp_data,PolCount,Value,1:21)

ggplot(data= temp2, aes (x = PolCount,y = Value,fill = Value)) + 
  geom_bar(stat = 'identity' )  +
  ggtitle("Insurance Policy Contribution") + 
  coord_flip()


```






### From the GBM variance influence chart, we will  select only 56  variables that have impact  The GBM variance influence chart was generated separately and will be  included at a later state

```{r}
Caravan_train_SUB56  <-  subset(Caravan_train, select = c(   car_policies   
                                                      ,Lower_level_education 
                                                      ,fire_policies                         
  ,Average_income                        
  ,private_third_party_insurance         
  ,X1_car                                
  ,Social_class_C                        
  ,Other_religion                        
  ,boat_policies                         
  ,No_car                                
  ,Customer_Subtype                      
  ,Middle_management                     
  ,Income_75.122.000                     
  ,tractor_policies                      
  ,Social_class_B1                       
  ,Social_class_D                        
  ,High_level_education                  
  ,Protestant                            
  ,Medium_level_education                
  ,bicycle_policies                      
  ,Income_45.75.000                      
  ,Unskilled_labourers                   
  ,Farmer                                
  ,Purchasing_power                      
  ,Number_of_car_policies                
  ,Income_._30                           
  ,Customer_main_type                    
  ,Household_with_children               
  ,Income_30.45.000                      
  ,Rented_house                          
  ,moped_policies                        
  ,Married                               
  ,Social_class_B2                       
  ,X2_cars                               
  ,Skilled_labourers                     
  ,Social_class_A                        
  ,Household_without_children            
  ,High_status                           
  ,No_religion                           
  ,Entrepreneur                          
  ,social_security_insurance_policies    
  ,Home_owners                           
  ,third_party_insurane_agriculture      
  ,Living_together                       
  ,Avg_size_household                    
  ,Avg_age                               
  ,Number_of_houses                      
  ,Singles                               
  ,Roman_catholic                        
  ,Other_relation                        
  ,Income_.123.000                       
  ,life_insurances                       
  ,National_Health_Service               
  ,Contribution_motorcycle_scooter       
  ,Private_health_insurance              
  ,Number_of_life_insurances  , Number_of_mobile_home_policies   )  )  

dim(Caravan_train_SUB56)
head(Caravan_train_SUB56)
                                                      
```


## CLASSIFICATION USING RANDOM  FOREST
###   STEP 1  - USE UNBALANCED DATA 
###   STEP 2  - USE UNDER SAMPLED DATA 
###   STEP 3 - PERFORMANCE TUNING 
   
### RF USING UNBALANCED DATA
```{r}
# use unbalanced data 



set.seed(123457)

attach(Caravan_train_SUB56)
rf.Unb.Train <-  randomForest( as.factor(Number_of_mobile_home_policies)  ~ . , data = Caravan_train_SUB56,importance = TRUE )
print(rf.Unb.Train)

# grab the error matrix 

err <- rf.Unb.Train$err.rate
head(err)
dim(rf.Unb.Train$err.rate)

# Look at final OOB error rate (last row in err matrix)
oob_err <- err[nrow(err),"OOB"]
print(oob_err)

# Plot the model trained in the previous exercise
plot(rf.Unb.Train)

# Add a legend since it doesn't have one by default
legend(x = "right", 
       legend = colnames(err),
       fill = 1:ncol(err))

#important(rf.Unb.Train)
varImpPlot(rf.Unb.Train)


CinsTest <-  read.csv.sql("C2.csv","select * from file where ORIGIN = 'test' ", header=TRUE, sep=",")


# Exclude the  first column 
# Caravan_test <- CinsTest[-1]

Caravan_test <-  subset(CinsTest, select = c(   car_policies   
                                                      ,Lower_level_education 
                                                      ,fire_policies                         
  ,Average_income                        
  ,private_third_party_insurance         
  ,X1_car                                
  ,Social_class_C                        
  ,Other_religion                        
  ,boat_policies                         
  ,No_car                                
  ,Customer_Subtype                      
  ,Middle_management                     
  ,Income_75.122.000                     
  ,tractor_policies                      
  ,Social_class_B1                       
  ,Social_class_D                        
  ,High_level_education                  
  ,Protestant                            
  ,Medium_level_education                
  ,bicycle_policies                      
  ,Income_45.75.000                      
  ,Unskilled_labourers                   
  ,Farmer                                
  ,Purchasing_power                      
  ,Number_of_car_policies                
  ,Income_._30                           
  ,Customer_main_type                    
  ,Household_with_children               
  ,Income_30.45.000                      
  ,Rented_house                          
  ,moped_policies                        
  ,Married                               
  ,Social_class_B2                       
  ,X2_cars                               
  ,Skilled_labourers                     
  ,Social_class_A                        
  ,Household_without_children            
  ,High_status                           
  ,No_religion                           
  ,Entrepreneur                          
  ,social_security_insurance_policies    
  ,Home_owners                           
  ,third_party_insurane_agriculture      
  ,Living_together                       
  ,Avg_size_household                    
  ,Avg_age                               
  ,Number_of_houses                      
  ,Singles                               
  ,Roman_catholic                        
  ,Other_relation                        
  ,Income_.123.000                       
  ,life_insurances                       
  ,National_Health_Service               
  ,Contribution_motorcycle_scooter       
  ,Private_health_insurance              
  ,Number_of_life_insurances  , Number_of_mobile_home_policies   )  )  

  
  
dim(Caravan_test)

names(Caravan_test)

# Generate predicted classes using the model object
rf.Unb.Prediction <-  predict(rf.Unb.Train,newdata = Caravan_test,type = "class")

# Calculate the confusion matrix for the test set
install.packages("caret")
install.packages('e1071', dependencies=TRUE)
library(caret)
cm <- confusionMatrix(as.factor(rf.Unb.Prediction),         
                      reference = as.factor(Caravan_test$Number_of_mobile_home_policies) )
print(cm)

# Compare test set accuracy to OOB accuracy
paste0("Test Accuracy: ", cm$overall[1])
paste0("OOB Accuracy: ", 1 - oob_err)


```

### RF USING UNDER SAMPLED DATA 
```{r}

attach(Caravan_train_SUB56)
caravan_under_data <-  ovun.sample( Number_of_mobile_home_policies ~ . , data = Caravan_train_SUB56 , method = "under",
                           N = 700 , seed = 1 )$data
table(caravan_under_data$Number_of_mobile_home_policies)



attach(caravan_under_data)
rf.Under.Samp.Train <-  randomForest( as.factor(Number_of_mobile_home_policies)  ~ . , data = caravan_under_data,importance = TRUE )
print(rf.Under.Samp.Train)

# grab the error matrix 

err <- rf.Under.Samp.Train$err.rate
head(err)
dim(rf.Under.Samp.Train$err.rate)

# Look at final OOB error rate (last row in err matrix)
oob_err <- err[nrow(err),"OOB"]
print(oob_err)

# Plot the model trained in the previous exercise
plot(rf.Under.Samp.Train)

# Add a legend since it doesn't have one by default
legend(x = "right", 
       legend = colnames(err),
       fill = 1:ncol(err))

#importan(rf.Under.Samp.Train)
varImpPlot(rf.Under.Samp.Train)



# Generate predicted classes using the model object
rf.Under.Samp.Prediction <-  predict(rf.Under.Samp.Train,newdata = Caravan_test,type = "class")

# Calculate the confusion matrix for the test set

cm <- confusionMatrix(as.factor(rf.Under.Samp.Prediction),         
                      reference = as.factor(Caravan_test$Number_of_mobile_home_policies) )
print(cm)

# Compare test set accuracy to OOB accuracy
paste0("Test Accuracy: ", cm$overall[1])
paste0("OOB Accuracy: ", 1 - oob_err)

# Compute the AUC (`actual` must be a binary 1/0 numeric vector)
#library(ROCR)
#install.packages("pROC")
#library(pROC)
#auc(Caravan_test$Number_of_mobile_home_policies,rf.Unb.Prediction)
```